---
title: "1 - HMD Dataset"
author: "Yiu Chung Lui (Gordon)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prerequisite - Install Necessary Packages

Install Packages (run if needed)
```{r, eval = FALSE}
install.packages(c("StMoMo",
                   "demography",
                   "stats",
                   "forecast",
                   "vars",
                   "MASS"))
```

## 1.1 - Obtain HMD dataset

To obtain the necessary datasets from HMD, we use the "demography" package, which requires a registered HMD account.

### Username and Password (insert HMD username and password here)
```{r}
username <- "" #Insert HMD account username
password <- "" #Insert HMD account password
```

### Load Packages
```{r}
library(StMoMo)
library(demography)
library(stats)
```

### HMD Countries

The HMD country codes and labels are organised using a CSV file. Users do not need to follow this exact approach, provided they supply the country codes in the same format. The labels are used only for display purposes and can be customised as desired.

```{r}
HMD_Countries <- read.csv("HMD_Countries.csv")
print(HMD_Countries$HMD_Country_Code)
print(HMD_Countries$Label)
```

### Filter the dataset to the selected ages and years (use ages 60-90 and years 1970 - 2019 in this example)
```{r, eval = FALSE}
year_range <- 1970:2019
age_lower <- 60
age_upper <- 90
```

### Create HMD Datasets
```{r, eval = FALSE}
# Define countries and labels
countries <- HMD_Countries$HMD_Country_Code
labels <- HMD_Countries$Label

# Loop and assign each dataset
for (i in 1:length(countries)) {
  # Obtain data from HMD
  HMD_data <- hmd.mx(country = countries[i],
                     username = username,
                     password = password,
                     label = labels[i])
  
  # Assign dataset name (e.g., UK_data, ENW_data)
  assign(paste0(labels[i], "_data"), HMD_data)
  
  # Extract year range
  HMD_sub <- extract.years(HMD_data, years = year_range)
  
  # Age filtering
  age_filter <- which(HMD_sub$age >= age_lower & HMD_sub$age <= age_upper)
  
  HMD_sub$rate$male   <- HMD_sub$rate$male[age_filter, ]
  HMD_sub$rate$female <- HMD_sub$rate$female[age_filter, ]
  HMD_sub$rate$total  <- HMD_sub$rate$total[age_filter, ]
  
  HMD_sub$pop$male    <- HMD_sub$pop$male[age_filter, ]
  HMD_sub$pop$female  <- HMD_sub$pop$female[age_filter, ]
  HMD_sub$pop$total   <- HMD_sub$pop$total[age_filter, ]
  
  HMD_sub$age <- HMD_sub$age[age_filter]
  
  # Assign subset object (e.g., UK_sub, ENW_sub)
  assign(paste0(labels[i], "_sub"), HMD_sub)
  
  # Convert into StMoMo Dataset (central exposures)
  stmomo_c <- StMoMoData(HMD_sub, series = "total", type = "central")
  assign(paste0(labels[i], "_stmomo_c"), stmomo_c)
  
  # Convert to initial exposures
  stmomo_i <- central2initial(stmomo_c)
  assign(paste0(labels[i], "_stmomo_i"), stmomo_i)
}
```

### Combine all population datasets
```{r, eval = FALSE}
# Loop over labels and return the corresponding datasets
stmomo_datasets <- lapply(labels, function(lbl) {
  get(paste0(lbl, "_stmomo_i"))
})

# Name the list by labels
names(stmomo_datasets) <- labels
```


## 1.2 - Create Alternative populations

In addition to directly obtaining population data from HMD, the following examples demonstrate how to construct additional datasets by combining death rates and exposure data from selected HMD populations. Methods such as simulating a population using a binomial model of deaths are also presented. These newly generated populations are then combined with the full HMD datasets obtained previously.

### Function to combine population datasets ("combine_stmomo")
```{r, eval = FALSE}
combine_stmomo <- function(countries, label) {
  # Obtain datasets with the specified countries
  dataset_list <- stmomo_datasets[countries]
  
  # Extract ages and years (all datasets share the same structure)
  ages <- dataset_list[[1]]$ages
  years <- dataset_list[[1]]$years
  
  # Empty combined matrices
  D_total <- matrix(0, nrow = length(ages), ncol = length(years),
                    dimnames = list(ages, years))
  E_total <- matrix(0, nrow = length(ages), ncol = length(years),
                    dimnames = list(ages, years))
  
  # Loop through each dataset and align before summing
  for (i in 1:length(dataset_list)) {
    # Ensure identical indexing
    ds_D <- dataset_list[[i]]$Dxt
    ds_E <- dataset_list[[i]]$Ext
    
    D_total <- D_total + ds_D
    E_total <- E_total + ds_E
  }
  
  # Build combined object
  stmomo_combined <- list(
    ages = ages,
    years = years,
    Dxt = D_total,
    Ext = E_total,
    type = "initial",
    series = "total",
    label = label
  )
  
  class(stmomo_combined) <- "StMoMoData"
  return(stmomo_combined)
}
```

### Combine populations (Europe example)
```{r, eval = FALSE}
# List Europe countries (Must be in stmomo_dataset)
Europe <- c("AUT", "BEL", "BGR", "CZE", "DNK", "FIN", "FRA", "HUN", "ISL", 
            "IRL", "LTU", "LUX", "NLD", "NOR", "POL", "PRT", "SVK", "ESP", 
            "SWE", "CHE", "UK")
# Also group Europe by geographical locations
Northern_Europe <- c("DNK","FIN","ISL","IRL","NOR","SWE","UK")
Western_Europe <- c("AUT","BEL","FRA","LUX","NLD","CHE")
Southern_Europe <- c("PRT","ESP")
Eastern_Europe <- c("BGR","CZE","HUN","LTU","POL","SVK")

# Use combine_stmomo function to create dataset for Europe Countries
Europe_stmomo_i <- combine_stmomo(Europe, label = "Europe")
Northern_Europe_stmomo_i <- combine_stmomo(Northern_Europe, label = "Northern_Europe")
Western_Europe_stmomo_i <- combine_stmomo(Western_Europe, label = "Western_Europe")
Southern_Europe_stmomo_i <- combine_stmomo(Southern_Europe, label = "Southern_Europe")
Eastern_Europe_stmomo_i <- combine_stmomo(Eastern_Europe , label = "Eastern_Europe ")

# Add Europe dataset to the existing list
stmomo_datasets$Europe <- Europe_stmomo_i
stmomo_datasets$Northern_Europe <- Northern_Europe_stmomo_i
stmomo_datasets$Western_Europe <- Western_Europe_stmomo_i
stmomo_datasets$Southern_Europe <- Southern_Europe_stmomo_i
stmomo_datasets$Eastern_Europe <- Eastern_Europe_stmomo_i
```

### Function to simulate populations ("simulate_country_population")
```{r, eval = FALSE}
simulate_country_population <- function(country, size_relative_to_country_population, set_seed = 1) {
  # Get country StMoMo dataset 
  data <- stmomo_datasets[[country]]
  data_qxt <- data$Dxt / data$Ext
  
  # Size of simulated population relative to the country population 
  size <- size_relative_to_country_population
  
  # Simulating B population
  Ext <- (data$Ext)/size
  Dxt <- Ext
  
  set.seed(set_seed)
  for (i in 1:(age_upper - age_lower + 1)) {
    for (j in 1:length(year_range)) {
      Dxt[i, j] <- rbinom(1, round(Ext[i, j], 0), data_qxt[i, j])
    }
  }
  
  # Form simulated population StMoMo Dataset
  sim_stmomo_i <- data
  sim_stmomo_i$label <- paste0("sim_", country)
  sim_stmomo_i$Ext <- Ext
  sim_stmomo_i$Dxt <- Dxt
  
  # Return StMoMo Dataset
  return(sim_stmomo_i)
}

```

### Simulating populations (England and Wales, Scotland, and Northern Ireland example)
```{r}
# Use simulate_country_population function to create dataset for simulated ENW, SCO, and NIR population
sim_ENW_stmomo_i <- simulate_country_population("ENW", size_relative_to_country_population = 5, set_seed = 1)
sim_SCO_stmomo_i <- simulate_country_population("SCO", size_relative_to_country_population = 3, set_seed = 1)
sim_NIR_stmomo_i <- simulate_country_population("NIR", size_relative_to_country_population = 2, set_seed = 1)
# set lower relative size due to small exposure

# Add simulated ENW, SCO, and NIR dataset to the existing list
stmomo_datasets$sim_ENW <- sim_ENW_stmomo_i
stmomo_datasets$sim_SCO <- sim_SCO_stmomo_i
stmomo_datasets$sim_NIR <- sim_NIR_stmomo_i
```

## Save Environment
```{r, eval = FALSE}
save.image(file = "environment.RData")
```

