---
title: "5 - Hedging"
author: "Yiu Chung Lui (Gordon)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prerequisite - Load Previous Environment
```{r}
load("environment.RData")
```

We evaluate index-based hedging instruments calibrated under the three-population mortality models. In particular, we construct a portfolio of index-based longevity swaps based on the two reference populations to hedge the book population’s mortality experience.

Longevity risk reduction is defined as:

```math
\text{Longevity Risk Reduction} = \left(1 - \frac{\text{Hedged Risk}}{\text{Unhedged Risk}}\right)\times 100\%.
```

For each simulation \(i\), the present value of the hedged portfolio is:

```math
PV^{(i)} =
\sum_{t=1}^{25} \frac{l^{B(i)}_{65+t,\,2020+t}}{(1+r)^t}
- w_1 \sum_{t=1}^{25} \frac{\left({}_tp^{R1(i)}_{65,2020} - {}_tp^{R1,\text{forward}}_{65,2020}\right)}{(1+r)^t}
- w_2 \sum_{t=1}^{25} \frac{\left({}_tp^{R2(i)}_{65,2020} - {}_tp^{R2,\text{forward}}_{65,2020}\right)}{(1+r)^t},
\quad i = 1,\ldots,5000.
```

### M7 equation function ("M7_logit_qxt")
```{r}
# M7 Equation
M7_logit_qxt <- function(kt, gc, age_range, year_range = NULL) {
  # Exception test for year_range
  if (is.null(year_range)) {
      year_range <- colnames(kt)
      if (is.null(year_range)) {
          stop("Need to provide 'year_range' or colnames(kt).")
      }
  }

  # Age terms
  ages <- age_range
  xbar <- mean(ages)
  ages_c  <- ages - xbar
  ages_c2 <- ages_c^2 - mean(ages_c^2)
  
  # Length of ages and years
  length_x <- length(age_range)
  length_t <- length(year_range)
  
  # Expand kt over ages (repeated rows)
  kt1 <- matrix(kt[1, ], nrow = length_x, ncol = length_t, byrow = TRUE)
  kt2 <- matrix(kt[2, ], nrow = length_x, ncol = length_t, byrow = TRUE)
  kt3 <- matrix(kt[3, ], nrow = length_x, ncol = length_t, byrow = TRUE)

  # Expand age terms over years (repeated columns)
  A1 <- matrix(ages_c,  nrow = length_x, ncol = length_t)
  A2 <- matrix(ages_c2, nrow = length_x, ncol = length_t)
  
  # Cohort term
  cohort_index <- outer(ages, as.numeric(year_range), function(x, t) t - x)
  gc_years <- as.numeric(labels(gc)[[2]])
  
  gc_mat <- matrix(NA, nrow = length_x, ncol = length_t)
  gc_mat[ , ] <- gc[match(cohort_index, gc_years)]
  
  # Calculate eta_xt
  eta_xt <- kt1 + kt2*A1 + kt3*A2 + gc_mat # M7 equation
  rownames(eta_xt) <- age_range
  colnames(eta_xt) <- year_range
  
  return(eta_xt)
}
```

### M6 equation function ("M6_logit_qxt")
```{r}
# M6 Equation
M6_logit_qxt <- function(kt, gc, age_range, year_range = NULL) {
  # Exception test for year_range
  if (is.null(year_range)) {
      year_range <- colnames(kt)
      if (is.null(year_range)) {
          stop("Need to provide 'year_range' or colnames(kt).")
      }
  }

  # Age terms
  ages <- age_range
  xbar <- mean(ages)
  ages_c  <- ages - xbar
  
  # Length of ages and years
  length_x <- length(age_range)
  length_t <- length(year_range)
  
  # Expand kt over ages (repeated rows)
  kt1 <- matrix(kt[1, ], nrow = length_x, ncol = length_t, byrow = TRUE)
  kt2 <- matrix(kt[2, ], nrow = length_x, ncol = length_t, byrow = TRUE)

  # Expand age terms over years (repeated columns)
  A1 <- matrix(ages_c,  nrow = length_x, ncol = length_t)
  
  # Cohort term
  cohort_index <- outer(ages, as.numeric(year_range), function(x, t) t - x)
  gc_years <- as.numeric(labels(gc)[[2]])
  
  gc_mat <- matrix(NA, nrow = length_x, ncol = length_t)
  gc_mat[ , ] <- gc[match(cohort_index, gc_years)]
  
  # Calculate eta_xt
  eta_xt <- kt1 + kt2*A1 + gc_mat # M6 equation
  rownames(eta_xt) <- age_range
  colnames(eta_xt) <- year_range
  
  return(eta_xt)
}
```

### M5 equation function ("M5_logit_qxt")
```{r}
# M6 Equation
M5_logit_qxt <- function(kt, age_range, year_range = NULL) {
  # Exception test for year_range
  if (is.null(year_range)) {
      year_range <- colnames(kt)
      if (is.null(year_range)) {
          stop("Need to provide 'year_range' or colnames(kt).")
      }
  }

  # Age terms
  ages <- age_range
  xbar <- mean(ages)
  ages_c  <- ages - xbar
  
  # Length of ages and years
  length_x <- length(age_range)
  length_t <- length(year_range)
  
  # Expand kt over ages (repeated rows)
  kt1 <- matrix(kt[1, ], nrow = length_x, ncol = length_t, byrow = TRUE)
  kt2 <- matrix(kt[2, ], nrow = length_x, ncol = length_t, byrow = TRUE)

  # Expand age terms over years (repeated columns)
  A1 <- matrix(ages_c,  nrow = length_x, ncol = length_t)
  
  # Calculate eta_xt
  eta_xt <- kt1 + kt2*A1 # M5 equation
  rownames(eta_xt) <- age_range
  colnames(eta_xt) <- year_range
  
  return(eta_xt)
}
```

### Pick the correct function for R1, R2 and B_3
```{r}
# Pick the correct function
R1_logit_qxt <- 
  if (identical(R1_model, M7)) {
    M7_logit_qxt
  } else if (identical(R1_model, M6)) {
    M6_logit_qxt
  } else if (identical(R1_model, M5)) {
    M5_logit_qxt
  } else {
    stop("R1 Model specification not included/correct")
  }

R2_logit_qxt <- 
  if (identical(R2_model, M7)) {
    M7_logit_qxt
  } else if (identical(R2_model, M6)) {
    M6_logit_qxt
  } else if (identical(R2_model, M5)) {
    M5_logit_qxt
  } else {
    stop("R2 Model specification not included/correct")
  }

B_3_logit_qxt <- 
  if (identical(B_3_model, M7)) {
    M7_logit_qxt
  } else if (identical(B_3_model, M6)) {
    M6_logit_qxt
  } else if (identical(B_3_model, M5)) {
    M5_logit_qxt
  } else {
    stop("B_3 Model specification not included/correct")
  }
```

### Calculate qxt for R1, R2 and B_3

### Obtain mean logit qxt
```{r}
age_range <- 60:90

# R1 logit qxt
if (identical(R1_model, M5)) {
  mean_R1_combined <- R1_logit_qxt(mean_kt_R1_combined, age_range = age_range)
} else {
  mean_R1_combined <- R1_logit_qxt(mean_kt_R1_combined, mean_gc_R1_combined, age_range = age_range)
}

# R2 logit qxt
if (identical(R2_model, M5)) {
  mean_R2_combined <- R2_logit_qxt(mean_kt_R2_combined, age_range = age_range) + mean_R1_combined
} else {
  mean_R2_combined <- R2_logit_qxt(mean_kt_R2_combined, mean_gc_R2_combined, age_range = age_range) + mean_R1_combined
}

# B_3 logit qxt
if (identical(B_3_model, M5)) {
  mean_B_3_combined <- B_3_logit_qxt(mean_kt_B_3_combined, age_range = age_range) + mean_R2_combined
} else {
  mean_B_3_combined <- B_3_logit_qxt(mean_kt_B_3_combined, mean_gc_B_3_combined, age_range = age_range) + mean_R2_combined
}
```

### Obtain simulated logit qxt
```{r}
sim_R1_combined <- vector("list", nBoot)
sim_R2_combined <- vector("list", nBoot)
sim_B_3_combined <- vector("list", nBoot)

for (i in 1:nBoot) {
  # R1 logit qxt
  if (identical(R1_model, M5)) {
    sim_R1_combined[[i]] <- R1_logit_qxt(sim_kt_R1_combined[[i]], age_range = age_range)
  } else {
    sim_R1_combined[[i]] <- R1_logit_qxt(sim_kt_R1_combined[[i]], sim_gc_R1_combined[[i]], age_range = age_range)
  }

  # R2 logit qxt
  if (identical(R2_model, M5)) {
    sim_R2_combined[[i]] <- R2_logit_qxt(sim_kt_R2_combined[[i]], age_range = age_range) + sim_R1_combined[[i]]
  } else {
    sim_R2_combined[[i]] <- R2_logit_qxt(sim_kt_R2_combined[[i]], sim_gc_R2_combined[[i]], age_range = age_range) + sim_R1_combined[[i]]
  }

  # B_3 logit qxt
  if (identical(B_3_model, M5)) {
    sim_B_3_combined[[i]] <- B_3_logit_qxt(sim_kt_B_3_combined[[i]], age_range = age_range) + sim_R2_combined[[i]]
  } else {
    sim_B_3_combined[[i]] <- B_3_logit_qxt(sim_kt_B_3_combined[[i]], sim_gc_B_3_combined[[i]], age_range = age_range) + sim_R2_combined[[i]]
  }
}
```

### Transform to qxt
```{r}
# Transform using the inv_logit function

# Mean qxt
mean_R1_qxt <- inv_logit(mean_R1_combined)
mean_R2_qxt <- inv_logit(mean_R2_combined)
mean_B_3_qxt <- inv_logit(mean_B_3_combined)

# Simulated qxt
sim_R1_qxt <- vector("list", nBoot)
sim_R2_qxt <- vector("list", nBoot)
sim_B_3_qxt <- vector("list", nBoot)

for (i in 1:nBoot) {
  sim_R1_qxt[[i]] <- inv_logit(sim_R1_combined[[i]])
  sim_R2_qxt[[i]] <- inv_logit(sim_R2_combined[[i]])
  sim_B_3_qxt[[i]] <- inv_logit(sim_B_3_combined[[i]])
}
```

### Calculate tpx (R1 and R2)

### tpx function ("tpx_function")
```{r}
tpx_function <- function(qxt_data, duration, starting_age, starting_year) {
  # Storing tpx
  tpx <- numeric(duration)
  names(tpx) <- paste0(1:duration, "p", starting_age)
  
  # Initializing tpx_dummy
  tpx_dummy <- 1
  
  # Loop to calculate tpx  
  for (t in 0:(duration - 1)) {
    # Current age and year
    age_x <- starting_age + t
    year_t <- starting_year + t
    
    # Extract qxt
    qxt <- qxt_data[as.character(age_x), as.character(year_t)]
    
    # Update survival probability
    tpx_dummy <- tpx_dummy * (1 - qxt)
    
    # Store tpx value
    tpx[t + 1] <- tpx_dummy
  }
  
  return(tpx)
}
```

### tPx forward (R1 and R2)
```{r}
tpx_R1_forward <- tpx_function(mean_R1_qxt, duration = 25, starting_age = 65, starting_year = 2020)
tpx_R2_forward <- tpx_function(mean_R2_qxt, duration = 25, starting_age = 65, starting_year = 2020)
```

### tpx simulated (R1 and R2)
```{r}
tpx_R1_simulated <- vector("list", nBoot)
tpx_R2_simulated <- vector("list", nBoot)

for (i in 1:nBoot) {
  tpx_R1_simulated[[i]] <- tpx_function(sim_R1_qxt[[i]], 
                                        duration = 25, 
                                        starting_age = 65, 
                                        starting_year = 2020)
  tpx_R2_simulated[[i]] <- tpx_function(sim_R2_qxt[[i]],
                                        duration = 25, 
                                        starting_age = 65, 
                                        starting_year = 2020)
}
```

### Calculate lx (B_3)

### lx function ("lx_function")
```{r}
lx_function <- function(qxt_data, duration, starting_age, starting_year, initial_portfolio_size, random_state = 1) {
  # Set seed
  set.seed(random_state)
  
  # Storing lx
  lx <- numeric(duration)
  names(lx) <- paste0("l", starting_age + (1:duration))
  
  # Initialize lx_dummy
  lx_dummy <- initial_portfolio_size
  
  # Loop to simulate lx
  for (t in 0:(duration - 1)) {
    # Current age and year
    age_x <- starting_age + t
    year_t <- starting_year + t
    
    # Extract qxt
    qxt <- qxt_data[as.character(age_x), as.character(year_t)]
    
    # Simulate and Update lx
    lx_dummy <- rbinom(n = 1, size = lx_dummy, prob = 1 - qxt)
    
    # Store tpx value
    lx[t + 1] <- lx_dummy
  }
  
  return(lx)
}
```

### lx_function example
```{r}
lx_function(mean_B_3_qxt, duration = 25, starting_age = 65, starting_year = 2020, 
            initial_portfolio_size = 30000, random_state = 1)
```

### Apply lx simulated on book
```{r}
lx_B_3_simulated <- vector("list", nBoot)

for (i in 1:nBoot) {
  lx_B_3_simulated[[i]] <- lx_function(sim_B_3_qxt[[i]], 
                                       duration = 25, 
                                       starting_age = 65, 
                                       starting_year = 2020,
                                       initial_portfolio_size = 30000,
                                       random_state = i) # Need to set random_state = i, can't be a constant
}
```

### Hedging Result (default is three-population)

```{r}
hedging_result <- function(lx_B, tpx_R1_f, tpx_R2_f, tpx_R1_sim, tpx_R2_sim, 
                           interest_rate, two_ref = TRUE, ref_pop = NULL) {
  # Defining parameters
  r <- interest_rate
  t <- seq(1, length(tpx_R1_f))
  
  # Set up matrix for A0, A1 and A2
  A0 <- numeric(nBoot) # Unhedged liability
  A1 <- numeric(nBoot) # R1 longevity swap
  A2 <- numeric(nBoot) # R2 longevity swap
  
  # Loop for each simulation
  for (i in 1:nBoot) {
    # Inputs for hedging equation
    A0[i] <- sum(lx_B[[i]]*(1 + r)^(-t))                    # Unhedged liability
    A1[i] <- sum((tpx_R1_sim[[i]] - tpx_R1_f)*(1 + r)^(-t)) # R1 longevity swap
    A2[i] <- sum((tpx_R2_sim[[i]] - tpx_R2_f)*(1 + r)^(-t)) # R2 longevity swap
  }
  
  # Define VaR equation
  VaR_demean <- function(payoff, p = 0.995) {
    unname(quantile(payoff - mean(payoff), probs = p))
  }
  
  # Define ES equation
  ES_demean <- function(payoff, p = 0.995) {
    # Demean payoff
    x <- payoff - mean(payoff)
  
    # Compute VaR threshold
    var_threshold <- unname(quantile(x, probs = p))
  
    # Compute Expected Shortfall Threshold
    mean(x[x >= var_threshold])
  }
  
  # Calculate unhedged var, sd, VaR and ES
  unhedged_var <- var(A0)
  unhedged_sd <- sqrt(unhedged_var)
  unhedged_VaR <- VaR_demean(A0)
  unhedged_ES <- ES_demean(A0)
  
  # Check if it uses two reference populations or one reference population
  if (two_ref == FALSE) {
    if (is.null(ref_pop) || !(ref_pop %in% c("R1","R2"))) {
      stop("If two_ref = FALSE, ref_pop must be 'R1' or 'R2'.")
    }
  }
  
  # Define Hedging equation
  hedging_eq <- function(w) {
    # Two reference populations
    if (two_ref) {
      w1 <- w[1]
      w2 <- w[2]
    } else {
      # One reference population
      if (ref_pop == "R1") {
        w1 <- w[1]
        w2 <- 0
      } else { # ref_pop == "R2"
        w1 <- 0
        w2 <- w[1]
      }
    }
    
    hedged_payoff <- A0 - w1*A1 - w2*A2
    VaR_demean(hedged_payoff)
  }
  
  # Initial guess
  if (two_ref == TRUE) {
    w0 <- c(0, 0)
    } else {
    w0 <- 0 # Only one hedging instrument to optimize
  }
  
  # Run optimisation (unconstrained)
  if (two_ref == TRUE) {
    opt <- optim(
      par = w0,
      fn = hedging_eq,
      method = "Nelder-Mead"
    )
#      method = "L-BFGS-B",
#      lower = c(-Inf, -Inf),
#      upper = c(Inf, Inf)
  } else {
    opt <- optim(
      par = w0,
      fn = hedging_eq,
      method = "Brent",
      lower = -max(A0)*5,
      upper = max(A0)*5
    )
  }

  
  # Extract optimized weights
  if (two_ref == TRUE) {
    w1 <- opt$par[1]
    w2 <- opt$par[2]
  } else {
    if (ref_pop == "R1") {
      w1 <- opt$par[1]
      w2 <- 0
      } else {
      w1 <- 0
      w2 <- opt$par[1]
    }
  }
  
  # Calculate hedged var, sd, VaR, and ES
  hedged_payoff <- A0 - w1*A1 - w2*A2
  hedged_var <- var(hedged_payoff)
  hedged_sd <- sqrt(hedged_var)
  hedged_VaR <- VaR_demean(hedged_payoff)
  hedged_ES <- ES_demean(hedged_payoff)
  
  # Calculate longevity risk reduction
  var_risk_reduction <- round((1 - hedged_var/unhedged_var)*100, 3)
  sd_risk_reduction <- round((1 - hedged_sd/unhedged_sd)*100, 3)
  VaR_risk_reduction <- round((1 - hedged_VaR/unhedged_VaR)*100, 3)
  ES_risk_reduction <- round((1 - hedged_ES/unhedged_ES)*100, 3)
  
  # Density plot
  
  # Combine x-range for both densities
  x_range <- range(c(A0 - mean(A0), hedged_payoff - mean(hedged_payoff)))

  # Compute density objects
  d_unhedged <- density(A0 - mean(A0))
  d_hedged <- density(hedged_payoff - mean(hedged_payoff))

  # Determine max y-value for plot limits
  max_height <- max(d_unhedged$y, d_hedged$y)

  # Plot unhedged density
  plot(d_unhedged,
       xlim = x_range*1.5,
       ylim = c(0, max_height*1.1),
       col = "red",
       lwd = 2,
       main = "Unhedged vs Hedged De-meaned Payoff Density Plot",
       xlab = "Payoff (demeaned)")

  # Add hedged density
  lines(d_hedged, col = "blue", lwd = 2)

  # Add legend
  legend("topright",
        legend = c("Unhedged", "Hedged"),
        col = c("red", "blue"),
        lwd = 2,
        bty = "n")
  
  # Return weights
  weights <- matrix(c(w1, w2),
                    nrow = 1,
                    ncol = 2)
  rownames(weights) <- "weights"
  colnames(weights) <- c(paste0("w1 (", model_populations[1], ")"), 
                         paste0("w2 (", model_populations[2], ")"))
  
  # Return hedging results
  results <- matrix(c(unhedged_var, unhedged_sd, unhedged_VaR, unhedged_ES,
                     hedged_var,   hedged_sd,   hedged_VaR,   hedged_ES,
                     var_risk_reduction, sd_risk_reduction, VaR_risk_reduction, ES_risk_reduction),
                    nrow = 4, 
                    ncol = 3)
  rownames(results) <- c("var", "sd", "VaR", "ES")
  colnames(results) <- c("unhedged", "hedged", "risk reduction (in %)")
  
  # Return both weights and hedging results
  return(list(
    weights = weights,
    results = results
  ))
}
```

### Evaluate hedging result
```{r}
two_ref_results <- hedging_result(lx_B_3_simulated, tpx_R1_forward, tpx_R2_forward,
                           tpx_R1_simulated, tpx_R2_simulated, interest_rate = 0.01) # Also prints the plot
print(two_ref_results$weights)
cat("\n")
print(two_ref_results$results)
```
```{r}
R1_ref_results <- hedging_result(lx_B_3_simulated, tpx_R1_forward, tpx_R2_forward,
                           tpx_R1_simulated, tpx_R2_simulated, interest_rate = 0.01,
                           two_ref = FALSE, ref_pop = "R1") # Also prints the plot
print(R1_ref_results$weights)
cat("\n")
print(R1_ref_results$results)
```
```{r}
R2_ref_results <- hedging_result(lx_B_3_simulated, tpx_R1_forward, tpx_R2_forward,
                           tpx_R1_simulated, tpx_R2_simulated, interest_rate = 0.01,
                           two_ref = FALSE, ref_pop = "R2") # Also prints the plot
print(R2_ref_results$weights)
cat("\n")
print(R2_ref_results$results)
```



